---
title: "COMP3020: Social web analytics - Group Project"
author: |
  - Williams Jung (22095912) – 25% 
  - Name (x) – 25% 
  - Name (x) – 25%
  - Name (x) – 25%
date: "2025-10-14"
output: html_document
---

# Cover Page

**Unit:** COMP3020 – Social Web Analytics\
**Session:** Spring 2025\
**Subject Coordinator:** Dr Gizem Intepe\
**Due Date:** Friday, 17 October 2025

## Group Members and Contributions

| Name          | Student ID | Contribution (%) |
|---------------|------------|------------------|
| Williams Jung | 22095912   | 25%              |
| Name          | x          | 25%              |
| Name          | x          | 25%              |
| Name          | x          | 25%              |

**Total Contribution:** 100%

Each group member contributed to data collection, text analysis, network graph sections, clustering, visualisations, hypothesis testing, overall editing, and report writing.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r authetication and load packages}
library("atrrr")
library("tm")
library("SnowballC") 
library("wordcloud")
library("dplyr")

ls("package:atrrr")

user <- "elaliza.bsky.social"
password = "4rhc-o5ku-fnce-6fin"

auth(user = user, password = password)

```

```{r q.1.1. Collect Posts}
posts1 <- search_post("Climate Change", limit = 300)
posts2 <- search_post("Elon Musk", limit = 300)
posts3 <- search_post("Christmas Carols", limit = 300)


# take posts and make into data frames that contain author handle, text, 
# indexed at, like count, reply count, langs

#extract relevant fields and create df
dataframe1 = function(allposts, keyword){
  data.frame(
    keyword = keyword,
    author_handle = allposts$author_handle,
    text = allposts$text,
    indexed_at = allposts$indexed_at,
    like_count = allposts$like_count,
    reply_count = allposts$reply_count,
    langs = sapply(allposts$langs, function(x) paste(x, collapse = ",")),
    stringsAsFactors = FALSE
  )
}

df1 = dataframe1(posts1, "Climate Change")
df2 = dataframe1(posts2, "Elon Musk")
df3 = dataframe1(posts3, "Christmas Carols")

combinedposts = rbind(df1, df2, df3) #combine all the df

#print summary display

cat("Total posts collected:", nrow(combinedposts))
cat("Climate Change posts:", nrow(df1))
cat("Elon Musk posts:", nrow(df2))
cat("Christmas Carols:", nrow(df3))

cat("Sample of collected data:")
print(head(combinedposts[, c("keyword", "author_handle", "like_count", "reply_count")],5))

```

```{r q.1.2. Word Frequency Visualisation}

#define function to clean text and generate word cloud for the 3 keywords
cleanCloud <- function(x,y){
  post.corpus <-  Corpus(VectorSource(x)) #create text corpus
  post.corpus <- tm_map(post.corpus, function(i) iconv(i, to='ASCII', sub= '')) #convert text to ASCII
  
  #remove all url (websites)
  removeURL = content_transformer(function(x) gsub("http[s]?://[[:alnum:]./]+", "", x))
  post.corpus = tm_map(post.corpus, removeURL)
  
  custom_stopwords= c(stopwords("english"), "amp", "tp") #you can add more
  
  post.corpus = tm_map(post.corpus, content_transformer(tolower)) #convert text to lowercase
  post.corpus = tm_map(post.corpus, removeNumbers) #remove numerics
  post.corpus = tm_map(post.corpus, removePunctuation) #remove punctuation marks
  post.corpus = tm_map(post.corpus, removeWords, custom_stopwords) #remove standard and custom stopwords
  post.corpus = tm_map(post.corpus, stripWhitespace) #remove extra white space
  post.corpus = tm_map(post.corpus, stemDocument, language = "english") #
  
  tdm = TermDocumentMatrix(post.corpus, control = list(wordLengths= c(3, Inf)))
  
  #remove empty documents after cleaning , as cleaning may remove all words!!!
  empties = which(colSums(as.matrix(tdm)) == 0)
  if (length(empties) > 0) {
    tdm = tdm[, -empties]
  }
  
  tdmw = weightTfIdf(tdm) #require tdm matrix, and apply Tf-idf weighting
  T = as.matrix(tdmw)
  
  freqsw = rowSums(T) #calculate frequencies
  
  wordcloud(names(freqsw),
            freqsw,
            random.order = FALSE,
            min.freq = 6,
            max.words = 200,
            scale = c(1, .5))
  title(y)
  
  return(list(tdm = tdm, tdmw = tdmw, empties = empties))

}
par(mfrow = c(1,3), mar = c(2,2,3,2))

cat("Creating word cloud 1: Climate Change")
result1 = cleanCloud(df1$text, "Climate Change")

cat("Creating word cloud 2: Elon Musk")
result2 = cleanCloud(df2$text, "Elon Musk")

cat("Creating word cloud 3: Christmas Carols")
result3 = cleanCloud(df3$text, "Christmas Carols")

```

```{r q1.3. Hypothesis Testing}

keywordA = "Climate Change"
keywordB = "Elon musk"

likesA = df1$like_count
likesB = df2$like_count

cat("Comparing average number of likes between keyword A and keyword B ")
cat("Keyword A: ", keywordA)
cat("Keyword B: ", keywordB)
cat("Sample size of A =", length(likesA), ",and B =", length(likesB))

##State hypothesis
# H0 = There is no difference in average number of likes between keyword A and keyword B.
# Ha = There is a significant difference in average number of likes between keyword A and keyword B 

#stats
cat("Keyword A: ", keywordA, "\n",
    "Mean: ", mean(likesA),
    "Standard Deviation: ", sd(likesA),
    "Median: ", median(likesA),
    "Min: ", min(likesA),
    "Max: ", max(likesA))

cat("Keyword B: ", keywordB, "\n",
    "Mean: ", mean(likesB),
    "Standard Deviation: ", sd(likesB),
    "Median: ", median(likesB),
    "Min: ", min(likesB),
    "Max: ", max(likesB))

#diagnosis test
#par(mfrow = c(1,3), mar = c(2,2,3,2))
hist(likesA, main = paste(keywordA, "Like distribution histogram"),
     xlab = "Like Count", col = "blue", breaks = 20)
abline(v = mean(likesA), col = "red")

hist(likesB, main = paste(keywordB, "Like distribution histogram"),
     xlab = "Like Count", col = "green", breaks = 20)
abline(v = mean(likesB), col = "red")

#MORE TESTS!!!!!!!

#t-test
t_test = t.test(likesA, likesB, var.equal = TRUE)
print(t_test)

```

```{r q2.1. build the user graph}

library("igraph")

climateUser <- search_user("Climate Change", limit = 300) 
head(climateUser$handle) # checking variable is established

userInfo <- get_user_info(unique(climateUser$handle))
names(userInfo)

pos <- which.max(userInfo$followers_count) #which.max to find user with highest follower count
seedClimateHandle <- userInfo$actor_handle[pos]
seedClimateHandle

friends <- get_follows(seedClimateHandle, limit = 50) #retrieve 1st degree friends
friendsName = friends$actor_handle
n <-  length(friendsName)

friendsOfFriends <- sapply(friendsName, get_follows, limit = 50) #retrieve 2nd degree friends
head(friendsOfFriends[[1]]$actor_handle)

#creating edge list for graph analysis
climate.r <-  rep(seedClimateHandle,n)
el.climate <- cbind(climate.r,friendsName)
el.climate

el.friends = list()
for (i in 1:n){
  if(!is.null(friendsOfFriends[[i]]) &&
     nrow(friendsOfFriends[[i]]) > 0 &&
     "actor_handle" %in% names(friendsOfFriends[[i]])) {
    name_rep = rep(friendsName[i], length(friendsOfFriends[[i]]$actor_handle))
    el.friends[[i]] = cbind(name_rep,friendsOfFriends[[i]]$actor_handle)
  }
}

el.friends = el.friends[!sapply(el.friends, is.null)]
el.more = do.call(rbind, el.friends)
el = rbind(el.climate,el.more)

dim(el)
cat("Total edges: ", nrow(el), "\n")
cat("Unique users: ", length(unique(c(el[,2]))))

```

```{r q2.2. visualising the graph}

#full graph
par(mfrow = c(1,1))

g = graph_from_edgelist(el, directed = TRUE)
plot(g, layout = layout.fruchterman.reingold,
     vertex.size = 1, edge.arrow.size = 0.1,
     vertex.label.cex = 0.1, vertex.color= "pink",
     vertex.shape="circle", vertex.label.color= "black")

#simplified subgraph

top_users <- names(sort(degree(g, mode = "all"), decreasing = TRUE))[1:20]
subg <- induced_subgraph(g, vids = top_users)

plot(
  subg,
  layout = layout.fruchterman.reingold(subg),
  vertex.size = degree(subg, mode = "all") * 0.8,
  vertex.label.cex = 0.7,
  edge.arrow.size = 0.3,
  vertex.color = "orange",
  vertex.label.color = "black",
  main = "Top 20 Most Connected Users"
)


```

```{r q2.3. Centrality Analysis}

#degree centrality 

order(degree(g), decreasing = TRUE)
centralIDs = order(degree(g), decreasing = TRUE)[1:5]
V(g)[centralIDs] #show most central vertices

for (i in 1:5){
  cat(sprintf("%d. %s (degree = %d)\n", i, 
              V(g)$name[centralIDs[i]], 
              degree(g)[centralIDs[i]])
)
}

#betweenness centrality 

order(betweenness(g), decreasing = TRUE)
betweenIDs = order(betweenness(g), decreasing = TRUE)[1:5]
V(g)[betweenIDs]

for (i in 1:5){
  cat(sprintf("%d. %s (degree = %d)\n", i, 
              V(g)$name[betweenIDs[i]], 
              degree(g)[betweenIDs[i]])
  )
}

#plot visualisation

par(mfrow = c(1,3), mar = c(2,2,3,2))

hist(degree(g), breaks = 30, col = "blue", border = "white",
     main = "Degree Distribution", xlab = "Degree", ylab = "Frequency")

# Betweenness distribution
hist(betweenness(g), breaks = 30, col = "red", border = "white",
     main = "Betweenness Distribution", xlab = "Betweenness", ylab = "Frequency")

# Degree vs Betweenness
plot(degree(g), betweenness(g), pch = 16, col = "yellow",
     main = "Degree vs Betweenness", xlab = "Degree", ylab = "Betweenness")

# In-degree vs Out-degree
plot(degree(g, mode = "in"), degree(g, mode = "out"), pch = 16, col = "green",
     main = "In vs Out Degree", xlab = "In-Degree", ylab = "Out-Degree")
abline(a = 0, b = 1, col = "red", lty = 2)

graph.density(g)

degree.distribution(g)

```

```{r q3.1}

user.behaviour <- userInfo %>%
  dplyr::select(followers_count, follows_count, posts_count)

user.behaviour <- na.omit(user.behaviour)

scaled.behaviour <- scale(user.behaviour)

withinss <- numeric(10)
for (k in 1:10) {
  km <- kmeans(scaled.behaviour, centers = k, nstart = 10)
  withinss[k] <- km$tot.withinss
}
plot(1:10, withinss, type="b", xlab="Number of clusters",
     ylab="Total within-cluster sum of squares",
     main="Elbow Method to choose k")

set.seed(123) # for reproducibility
k3 <- kmeans(scaled.behaviour, centers = 3, nstart = 20)

userInfo$cluster <- k3$cluster

pca <- prcomp(scaled.behaviour)
summary(pca)

plot(pca$x[,1], pca$x[,2], col = userInfo$cluster,
     pch = 19, xlab = "PC1", ylab = "PC2",
     main = "User Clusters by Behavioural Features (PCA)")
legend("topright", legend = paste("Cluster", 1:3), col = 1:3, pch = 19)

```

```{r q3.2}

graph_handles <- unique(c(seedClimateHandle, friendsName))

# Initialize list to store posts
user_posts <- list()

# Retrieve posts for each user
for (i in 1:length(graph_handles)) {
  handle <- graph_handles[i]
  cat("Retrieving posts for user", i, "of", length(graph_handles), ":", handle, "\n")
  
  # Get up to 10 most recent posts
  posts <- get_skeets_authored_by(handle, limit = 10, filter = "posts_with_replies")
  
  # Store the posts
  user_posts[[handle]] <- posts
  
  # Brief pause to be respectful to the API
  if(i%%10==0) cat(sprintf("Progress: %d/%d\n", i, length(graph_handles)))
  Sys.sleep(0.2)
}


# Concatenate the posts into a single document per user to create one 
#row per user. (If you struggle with merging posts, 
#you may retrieve and use only the most recent post per user instead.)

user_documents = lapply(user_posts, function(posts) 
  if (!is.null(posts$text) && length(posts$text) > 0) {
    paste(posts$text, collapse = " ")
  })


# Build a TF-IDF matrix from the user documents.
# make a corpus
corpus = Corpus(VectorSource(user_documents))

corpus_clean = corpus %>% #clean text
  tm_map(content_transformer(tolower)) %>%
  tm_map(removeNumbers) %>%
  tm_map(removePunctuation) %>%
  tm_map(removeWords, stopwords("english")) %>%
  tm_map(stripWhitespace)


# Create TF-IDF matrix

tfidf_matrix = DocumentTermMatrix(corpus_clean, 
                                  control = list(weighting = weightTfIdf))

# Apply clustering (K-means) and visualise.

# Convert TF-IDF matrix to regular matrix for clustering
tfidf_reg = as.matrix(tfidf_matrix)

# elbow method to choose k
withinss = numeric(10)
for (k in 1:10) {
  km = kmeans(tfidf_reg, centers = k, nstart = 10)
  withinss[k] <- km$tot.withinss
}

plot(1:10, withinss, type = "b", 
     xlab = "Number of clusters",
     ylab = "Total within cluster sum of squares",
     main = "Elbow Method to find k means")

# use k3
set.seed(123)
k3_text = kmeans(tfidf_reg, centers = 3, nstart = 20)

text_clusters = data.frame(
  actor_handle = names(user_documents),
  text_cluster = k3_text$cluster,
  stringsAsFactors = FALSE
)

userInfo = userInfo %>%
  left_join(text_clusters, by = "actor_handle")

# Visualaise:

pca_text = prcomp(tfidf_reg, scale. = TRUE)

plot(pca_text$x[,1], pca_text$x[,2], 
     col = userInfo$text_cluster,
     pch = 19, xlab = "PC1", ylab = "PC2",
     main = "User Clusters by Text Content (TF-IDF PCA)")
legend("topright", legend = paste("Cluster", 1:3), col = 1:3, pch = 19)

```
